---
title: "Diamonds Prices"
date: today
author: Chu,Li,Hsu
format:
 pdf:
    engine: xelatex
    documentclass: article
    geometry: "left=2cm, right=2cm, top=2cm, bottom=2cm, a4paper"
    fontsize: 12pt
    header-includes:
      - \usepackage{xeCJK}
      - \setCJKmainfont{Microsoft JhengHei UI}
toc: true  
---

```{r}
library(showtext)
showtext_auto()  # 啟用 showtext
font_add("Microsoft JhengHei UI", "C:/Windows/Fonts/msjh.ttc")  # 添加你使用的字體
```

```{r library}
library(Hmisc)
library(skimr)
library(DataExplorer)
library(ggplot2)
library(dplyr)
library(corrplot)
library(GGally)
library(plotly)
library(gridExtra)
library(knitr)
library(car)
setwd("C:/Users/User/OneDrive/桌面/統諮期末")
data <- read.csv("Sleep_health_and_lifestyle_dataset.csv")
```

# 1. Conduct necessary data preprocessing

## 敘述性統計/missing values診斷
```{r descriptive statistics}
# Check structure of the dataset
head(data)
dim(data)
names(data)
str(data)      
skim(data)
describe(data)
sum(is.na(data))
plot_missing(data)
```

此筆資料集共有374筆資料，13個變數且無缺失值

其中gender,occupation,quality.of.sleep為類別變數;

age,sleep.duration,blood.pressure為連續變數




## Scaling for predicting
```{r scale}
# Scale numerical variables
#num_cols <- c("carat", "depth", "table", "price", "x", "y", "z")
#diamond[num_cols] <- scale(diamond[num_cols])
```

## table one
```{r table one}
summary_table <- data %>%
  summarise(
    Variable = c(
      "Person ID",
      "Gender",
      "Age",
      "Occupation",
      "Sleep Duration",
      "Quality of Sleep",
      "Physical Activity Level",
      "Stress Level",
      "BMI Category",
      "Blood Pressure",
      "Heart Rate",
      "Daily Steps",
      "Sleep Disorder"
    ),
    Description = c(
      "編號",
      "性別",
      "年齡",
      "職業",
      "每日睡眠時長(小時)",
      "主觀認定之睡眠品質",
      "Physical Activity Level",
      "主觀認定之壓力程度",
      "BMI類別",
      "血壓",
      "脈搏",
      "每日步數",
      "睡眠疾病"
    ),
    remark=c(
      "1-374",
      "Male/Female",
      "27-59",
      "Occupation",
      "Sleep Duration",
      "scale: 1-10",
      "Physical Activity Level",
      "scale: 1-10",
      "Underweight/Normal/Overweight...",
      "systolic/diastolic",
      "bpm",
      "Daily Steps",
      "None/Insomnia/Apnea"
    )
  )
kable(summary_table, format = "markdown", digits = 2, caption = " 變數解釋")

```

## 資料前處理
```{r pre-processing}
#刪除Person ID
data <- data %>% select(-`Person.ID`)

#把blood pressure分成兩col
data <- data %>%
  tidyr::separate(col = `Blood.Pressure`, 
                  into = c("BloodPressure_Upper", "BloodPressure_Lower"), 
                  sep = "/", 
                  convert = TRUE) # convert=TRUE 會自動轉換為數值型別
data$BloodPressure_Upper <- as.numeric(data$BloodPressure_Upper)
data$BloodPressure_Lower <- as.numeric(data$BloodPressure_Lower)

#分類physical activity level
data$Physical.Activity.Level<-ifelse(data$Physical.Activity.Level<=35,"<=35",
                              ifelse(data$Physical.Activity.Level<=45,"<=45",
                              ifelse(data$Physical.Activity.Level<=60,"<=60",
                              ifelse(data$Physical.Activity.Level<=75,"<=75",
                              "<=90"))))
#分類daily steps
data$Daily.Steps <- ifelse(data$Daily.Steps<=5000,"<=5000",
                    ifelse(data$Daily.Steps<=6000,"<=6000",
                    ifelse(data$Daily.Steps<=7000,"<=7000","7000up")))  

#將睡眠疾病->0,1
data$Sleep.Disorder <- ifelse(data$Sleep.Disorder=="None",0,1)

#分類BMI
data$BMI.Category <- ifelse(data$BMI.Category == "Normal Weight","Normal",
                            data$BMI.Category)
data$BMI.Category <- ifelse(data$BMI.Category == "Obese","Overweight",
                            data$BMI.Category)

#分類quality of sleep
data$Quality.of.Sleep <- ifelse(data$Quality.of.Sleep==4 | 
                                data$Quality.of.Sleep==5,"4-5",
                                data$Quality.of.Sleep)

#分類occupation
data$Occupation <- ifelse(data$Occupation=="Manager" | data$Occupation=="Sales Representative" ,"Salesperson",data$Occupation)
data$Occupation <- ifelse(data$Occupation=="Software Engineer"  ,"Engineer",data$Occupation)
```

## Encoding Categorical Variables
```{r encoding}
data$Gender <- as.factor(data$Gender)
data$Occupation <- as.factor(data$Occupation)
data$Quality.of.Sleep <- as.factor(data$Quality.of.Sleep)
data$Stress.Level <- as.factor(data$Stress.Level)
data$BMI.Category <- as.factor(data$BMI.Category)
data$Sleep.Disorder <- as.factor(data$Sleep.Disorder)
data$Physical.Activity.Level <- as.factor(data$Physical.Activity.Level)
data$Daily.Steps <- as.factor(data$Daily.Steps)
```

## 處理後的資料
```{r after}
describe(data)
```

描述性統計:比較不同組別間的變數分布差異
```{r descriptive2}
library(Hmisc) 
output0 <- summaryM(Age + Gender + Occupation + Sleep.Duration + 
             Quality.of.Sleep + Physical.Activity.Level + Stress.Level + 
             BMI.Category + BloodPressure_Upper + BloodPressure_Lower + 
             Heart.Rate + Daily.Steps
                                 ~ Sleep.Disorder,
                                 #~ 1,
                                 data = data, test = F, overall = F, na.include=T)
#sink(paste0("Table1.txt"))
print(output0, long=TRUE, what = "%")
```


# 2. EDA

## Distribution of the data

## i.categorical variable
```{r plot1}
p1 <- ggplot(data, aes(x = Gender, fill = Gender)) +
  geom_bar() +
  labs(title = "Count of Gender", x = "Gender", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3")
p2 <- ggplot(data, aes(x = Occupation, fill = Occupation)) +
  geom_bar() +
  labs(title = "Count of Occupation", x = "Occupation", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3")
p3 <- ggplot(data, aes(x = Quality.of.Sleep, fill = Quality.of.Sleep)) +
  geom_bar() +
  labs(title = "Count of Quality.of.Sleep", x = "Quality.of.Sleep", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3")
p4 <- ggplot(data, 
  aes(x = Physical.Activity.Level, fill = Physical.Activity.Level)) +
  geom_bar() +
  labs(title = "Count of Physical.Activity.Level",
  x = "Physical.Activity.Level", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3")
grid.arrange(p1,p2,p3,p4,ncol = 2)
p5 <- ggplot(data, aes(x = Stress.Level, fill = Stress.Level)) +
  geom_bar() +
  labs(title = "Count of Stress.Level", x = "Stress.Level", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3")
p6 <- ggplot(data, aes(x = BMI.Category, fill = BMI.Category)) +
  geom_bar() +
  labs(title = "Count of BMI.Category", x = "BMI.Category", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3")
p7 <- ggplot(data, aes(x = Daily.Steps, fill = Daily.Steps)) +
  geom_bar() +
  labs(title = "Count of Daily.Steps", x = "Daily.Steps", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3")
grid.arrange(p5,p6,p7,ncol = 2)
```

## ii.continuous variable
```{r,eval=FALSE}
layout(mat = matrix(c(1,2),2, byrow = FALSE),  height = c(8,1))
par(mar=c(4, 4, 3, 2))
hist(data$Age, main = 'Distribution of Age',
     xlab="Age",col="lightblue")
par(mar=c(0.5, 4, 0.5, 2))
boxplot(data$Age, xaxt = "n", horizontal=TRUE,  
        col="pink", border="black", frame = FALSE)

par(mar=c(4, 4, 3, 2))
hist(data$Sleep.Duration, main = 'Distribution of Sleep.Duration',
     xlab="Sleep.Duration",col="lightblue")
par(mar=c(0.5, 4, 0.5, 2))
boxplot(data$Sleep.Duration, xaxt = "n", horizontal=TRUE,  
        col="pink", border="black", frame = FALSE)

par(mar=c(4, 4, 3, 2))
hist(data$BloodPressure_Upper, main = 'Distribution of BloodPressure_Upper',
     xlab="BloodPressure_Upper",col="lightblue")
par(mar=c(0.5, 4, 0.5, 2))
boxplot(data$BloodPressure_Upper, xaxt = "n", horizontal=TRUE,  
        col="pink", border="black", frame = FALSE)

par(mar=c(4, 4, 3, 2))
hist(data$BloodPressure_Lower, main = 'Distribution of BloodPressure_Lower',
     xlab="BloodPressure_Lower",col="lightblue")
par(mar=c(0.5, 4, 0.5, 2))
boxplot(data$BloodPressure_Lower, xaxt = "n", horizontal=TRUE,  
        col="pink", border="black", frame = FALSE)

par(mar=c(4, 4, 3, 2))
hist(data$Heart.Rate, main = 'Distribution of Heart.Rate',
     xlab="Heart.Rate",col="lightblue")
par(mar=c(0.5, 4, 0.5, 2))
boxplot(data$Heart.Rate, xaxt = "n", horizontal=TRUE,  
        col="pink", border="black", frame = FALSE)
```

## iii.Sleep Disorder
```{r}
ggplot(data, aes(x = Sleep.Disorder, fill = Sleep.Disorder)) +
  geom_bar() +
  labs(title = "Count of Sleep.Disorder", x = "Sleep.Disorder", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set2")

data %>%
  group_by(`Sleep.Disorder`, Gender) %>%
  summarise(count = n(), .groups = "drop") 
```

## Correlation between data(variables & sleep disorder)

## i.categorical variable
```{r}
ggplot(data, aes(x = Gender, fill = Sleep.Disorder)) +
  geom_bar(position = "dodge") +
  labs(title = "Relationship between Gender and Sleep Disorder", 
       x = "Gender", 
       y = "Count") +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal()

ggplot(data, aes(x = Occupation, fill = Sleep.Disorder)) +
  geom_bar(position = "dodge") +
  labs(title = "Relationship between Occupation and Sleep Disorder", 
       x = "Occupation", 
       y = "Count") +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal()

ggplot(data, aes(x = Quality.of.Sleep, fill = Sleep.Disorder)) +
  geom_bar(position = "dodge") +
  labs(title = "Relationship between Sleep Quality and Sleep Disorder", 
       x = "Quality.of.Sleep", 
       y = "Count") +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal()

ggplot(data, aes(x = Physical.Activity.Level, fill = Sleep.Disorder)) +
  geom_bar(position = "dodge") +
  labs(title = "Relationship between Physical.Activity.Level and Sleep Disorder", 
       x = "Physical.Activity.Level", 
       y = "Count") +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal()

ggplot(data, aes(x = Stress.Level, fill = Sleep.Disorder)) +
  geom_bar(position = "dodge") +
  labs(title = "Relationship between Stress.Level and Sleep Disorder", 
       x = "Stress.Level", 
       y = "Count") +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal()

ggplot(data, aes(x = BMI.Category, fill = Sleep.Disorder)) +
  geom_bar(position = "dodge") +
  labs(title = "Relationship between BMI.Category and Sleep Disorder", 
       x = "BMI.Category", 
       y = "Count") +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal()

ggplot(data, aes(x = Daily.Steps, fill = Sleep.Disorder)) +
  geom_bar(position = "dodge") +
  labs(title = "Relationship between Daily.Steps and Sleep Disorder", 
       x = "Daily.Steps", 
       y = "Count") +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal()
```

## 卡方檢定
```{r chisq}
chisq.test(table(data$Occupation, data$Sleep.Disorder))
fisher.test(table(data$Occupation, data$Sleep.Disorder),simulate.p.value=TRUE)
library(vcd)
# 計算 Cramér's V
assocstats(table(data$Gender, data$Sleep.Disorder))$cramer
```

### 偷放幾個酷酷的圖
```{r cool}
library(ggmosaic)
# 繪製馬賽克圖
ggplot(data) +
  geom_mosaic(aes(x = product(Gender), fill = Sleep.Disorder)) +
  labs(title = "Mosaic Plot of Gender and Sleep Disorder", 
       x = "Gender", 
       y = "Proportion") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set2")

library(reshape2)
# 創建交叉表
table_data <- table(data$Gender, data$Sleep.Disorder)
heatmap_data <- as.data.frame(as.table(table_data))
# 繪製熱圖
ggplot(heatmap_data, aes(x = Var1, y = Var2, fill = Freq)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Freq), color = "black") +
  scale_fill_gradient(low = "lightblue", high = "red") +
  labs(title = "Heatmap of Gender and Sleep Disorder", 
       x = "Gender", 
       y = "Sleep Disorder") +
  theme_minimal()
```

## ii.continuous variable
```{r continuous1}
ggplot(data, aes(x = Sleep.Disorder, y = Age, fill = Sleep.Disorder)) +
  geom_boxplot() +
  labs(title = "Sleep disorder Distribution by Age", 
       x = "Sleep.Disorder", y = "Age") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set2")

ggplot(data, aes(x = Sleep.Disorder, y = Sleep.Duration, fill = Sleep.Disorder)) +
  geom_boxplot() +
  labs(title = "Sleep disorder Distribution by Sleep.Duration", 
       x = "Sleep.Disorder", y = "Sleep.Duration") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set2")

ggplot(data, aes(x = Sleep.Disorder, y = BloodPressure_Upper, fill = Sleep.Disorder)) +
  geom_boxplot() +
  labs(title = "Sleep disorder Distribution by BloodPressure_Upper", 
       x = "Sleep.Disorder", y = "BloodPressure_Upper") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set2")

ggplot(data, aes(x = Sleep.Disorder, y = BloodPressure_Lower, fill = Sleep.Disorder)) +
  geom_boxplot() +
  labs(title = "Sleep disorder Distribution by BloodPressure_Lower", 
       x = "Sleep.Disorder", y = "BloodPressure_Lower") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set2")
```

## 連續型自變數之間的關係
```{r correlation}
#heatmap
par(mfrow = c(1,1))
numeric_vars <- data %>% select(Age, Sleep.Duration, BloodPressure_Upper, BloodPressure_Lower)
cor_matrix <- cor(numeric_vars)
corrplot(cor_matrix, method = "number", type = "upper", 
         tl.col = "black", tl.srt = 45, 
         col = colorRampPalette(c("blue", "white", "red"))(200))
```

blood pressure間呈高度正相關，

變數間呈現負相關的組合:blood pressure & sleep duration

## 一些類別變數交互作用的圖

```{r,eval=FALSE}
ggpairs(data, aes(color = Sleep.Disorder, alpha = 0.6))

```

```{r,eval=FALSE}
p <- ggplot(data, aes(x = BloodPressure_Lower, y = Occupation, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Bloodpressure and Occupation",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
p <- ggplot(data, aes(x = BMI.Category, y = Occupation, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between BMI and Occupation",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
p <- ggplot(data, aes(x = Sleep.Duration, y = BMI.Category, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Sleep.Duration and BMI",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
p <- ggplot(data, aes(x = Sleep.Duration, y = Occupation, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Sleep.Duration and Occupation",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
p <- ggplot(data, aes(x = BloodPressure_Lower, y = Age, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Bloodpressure and Age",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
p <- ggplot(data, aes(x = Age, y = Occupation, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Age and Occupation",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
p <- ggplot(data, aes(x = Physical.Activity.Level, y = Occupation, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Physical.Activity.Level and Occupation",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
p <- ggplot(data, aes(x = Physical.Activity.Level, y = Daily.Steps, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Physical.Activity.Level and Daily.Steps",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
p <- ggplot(data, aes(x = Quality.of.Sleep, y = Occupation, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Quality.of.Sleep and Occupation",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
p <- ggplot(data, aes(x = Stress.Level, y = Quality.of.Sleep, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Stress.Level and Quality of Sleep",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
p <- ggplot(data, aes(x = Age, y = Quality.of.Sleep, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Age and Quality",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
p <- ggplot(data, aes(x = Physical.Activity.Level, y = Stress.Level, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Physical.Activity and Stress.level",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
p <- ggplot(data, aes(x = Stress.Level, y = Daily.Steps, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Stress.Level and Daily.Steps",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
p <- ggplot(data, aes(x = Stress.Level, y = Heart.Rate, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Stress.Level and Heart.Rate",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
p <- ggplot(data, aes(x = Stress.Level, y = Sleep.Duration, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Stress.Level and Sleep Duration",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
p <- ggplot(data, aes(x = Stress.Level, y = Age, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Stress.Level and Age",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
p <- ggplot(data, aes(x = Stress.Level, y = BloodPressure_Lower, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Stress.Level and BloodPressure_Lower",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
p <- ggplot(data, aes(x = Age, y = Physical.Activity.Level, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Age and Physical.Activity.Level",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
p <- ggplot(data, aes(x = BloodPressure_Lower, y = BMI.Category, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between BloodPressure_Lower and BMI",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
p <- ggplot(data, aes(x = Daily.Steps, y = Occupation, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Occupation and Daily.Steps",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
# 靜態氣泡圖
p <- ggplot(data, aes(x = Heart.Rate, y = Daily.Steps, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Heart.Rate and Daily.Steps",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
library(ggplot2)
library(plotly)

# 靜態氣泡圖
p <- ggplot(data, aes(x = Heart.Rate, y = BMI.Category, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Heart.Rate and BMI",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
library(ggplot2)
library(plotly)

# 靜態氣泡圖
p <- ggplot(data, aes(x = Stress.Level, y = BMI.Category, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Stress.Level and BMI",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot
```

```{r,eval=FALSE}
library(ggplot2)
library(plotly)

# 靜態氣泡圖
p <- ggplot(data, aes(x = Stress.Level, y = Occupation, color = Sleep.Disorder)) +
  geom_count() +
  scale_size_area(max_size = 10) +
  labs(title = "Interaction between Stress.Level and Occupation",
       size = "Count") +
  theme_minimal()

# 轉為交互式氣泡圖
interactive_plot <- ggplotly(p)

# 顯示交互式圖
interactive_plot

```

## 大整理:變數之間correlation計算(不同類型:連續vs.連續、類別vs.類別、類別vs.連續) 輸出excel檔

```{r summary}
# 提取變數名稱
all_vars <- names(data)

# 確定類別與連續變數
categorical_vars <- all_vars[sapply(data, is.factor)]
continuous_vars <- all_vars[sapply(data, is.numeric)]

# 初始化結果數據框
results <- data.frame(
  Variable1 = character(),
  Variable2 = character(),
  Correlation_Type = character(),
  Correlation_Value = numeric(),
  P_Value = numeric(),
  stringsAsFactors = FALSE
)

# 計算相關性
for (i in 1:(length(all_vars) - 1)) {
  for (j in (i + 1):length(all_vars)) {
    var1 <- all_vars[i]
    var2 <- all_vars[j]
    
    # 類別對類別
    if (var1 %in% categorical_vars && var2 %in% categorical_vars) {
      tbl <- table(data[[var1]], data[[var2]])
      chi_test <- chisq.test(tbl)
      n <- sum(tbl)
      min_dim <- min(nrow(tbl) - 1, ncol(tbl) - 1)
      cramers_v <- sqrt(chi_test$statistic / (n * min_dim))
      results <- rbind(results, data.frame(
        Variable1 = var1,
        Variable2 = var2,
        Correlation_Type = "Cramer's V",
        Correlation_Value = cramers_v,
        P_Value = chi_test$p.value
      ))
    
    # 類別對連續 (點二列相關)
    } else if ((var1 %in% categorical_vars && var2 %in% continuous_vars) ||
               (var1 %in% continuous_vars && var2 %in% categorical_vars)) {
      cat_var <- ifelse(var1 %in% categorical_vars, var1, var2)
      cont_var <- ifelse(var1 %in% continuous_vars, var1, var2)
      cor_test <- cor.test(as.numeric(data[[cat_var]]), data[[cont_var]])
      results <- rbind(results, data.frame(
        Variable1 = var1,
        Variable2 = var2,
        Correlation_Type = "Point-Biserial",
        Correlation_Value = cor_test$estimate,
        P_Value = cor_test$p.value
      ))
    
    # 連續對連續 (皮爾森相關)
    } else if (var1 %in% continuous_vars && var2 %in% continuous_vars) {
      cor_test <- cor.test(data[[var1]], data[[var2]])
      results <- rbind(results, data.frame(
        Variable1 = var1,
        Variable2 = var2,
        Correlation_Type = "Pearson",
        Correlation_Value = cor_test$estimate,
        P_Value = cor_test$p.value
      ))
    }
  }
}

# 將結果輸出為 CSV 文件
write.csv(results, "AllVariableCorrelationResults.csv", row.names = FALSE)

```

# 3. Construct a predictive model for sleep disorder
```{r library2}
library(caret)          # For data partitioning and confusion matrix
library(ROCR)           # For ROC curve and AUC
library(pROC) 
library(randomForest) 
library(xgboost)
library(Matrix)
library(pscl)
library(glmnet)
```

```{r}
set.seed(123) 
train_index <- createDataPartition(data$Sleep.Disorder, p = 0.8, list = FALSE)
train_data <- data[train_index, ]
test_data <- data[-train_index, ]
```

## logistic regression(全放/共線性非常高)
```{r}
model <- glm(Sleep.Disorder ~ Age + Gender + Occupation + Sleep.Duration + 
             Quality.of.Sleep + Physical.Activity.Level + Stress.Level + 
             BMI.Category + BloodPressure_Upper + BloodPressure_Lower + 
             Heart.Rate + Daily.Steps, 
             data = train_data, family = binomial())

summary(model)  

predicted_probabilities <- predict(model, newdata = test_data, type = "response")
predicted_classes <- ifelse(predicted_probabilities > 0.5, 1, 0)

# Confusion Matrix
confusion_matrix <- confusionMatrix(as.factor(predicted_classes), test_data$Sleep.Disorder)
print(confusion_matrix)
# ROC
roc_curve <- roc(test_data$Sleep.Disorder, predicted_probabilities)
plot(roc_curve, main = "ROC Curve for Sleep Disorder Prediction")
auc_value <- auc(roc_curve)
print(paste("AUC:", auc_value))
vif(model)
```

## logistic regression(stepwise挑變數/共線性還是有點高)

Sleep.Duration + Quality.of.Sleep + Physical.Activity.Level + 
Stress.Level + BloodPressure_Lower + Daily.Steps
```{r}
library(MASS)
model <- glm(Sleep.Disorder ~ Age + Gender + Occupation + Sleep.Duration + 
             Quality.of.Sleep + Physical.Activity.Level + Stress.Level + 
             BMI.Category + BloodPressure_Upper + BloodPressure_Lower + 
             Heart.Rate + Daily.Steps, 
             data = train_data, family = binomial())

logistic_model_step <- stepAIC(model, direction = "both")
summary(logistic_model_step)
vif(logistic_model_step)
pseudo_r2 <- pR2(logistic_model_step)
print(pseudo_r2)
predicted_probs <- predict(logistic_model_step, newdata=test_data,type = "response")
predicted_classes <- ifelse(predicted_probs > 0.4, 1, 0)
library(caret)
conf_matrix <- confusionMatrix(as.factor(predicted_classes), as.factor(test_data$Sleep.Disorder))
print(conf_matrix)
```

## logistic regression(Elastic net/共線性還是有點高)

```{r}
library(glmnet)

# 訓練 Elastic Net 模型
variablenames <- names(data)[-c(13:16)]
formula.x <- formula(paste("~", paste(variablenames, collapse=" + ")))
X <- model.matrix(formula.x, data)
y <- data$Sleep.Disorder

## Using cross validation folds to select lambda.
cv <- cv.glmnet(x=X, y=y, family = "binomial",  alpha = 0.5) ## alpha = 1, LASSO; = 0, ridge
coefs <- coef(cv, s=cv$lambda.1se)  
best_lambda <- cv$lambda.min
print(best_lambda)
fre.variables <- names(coefs[which(coefs[,1]!=0),1])
fre.variables
```

```{r}
logistic_model_select <- glm(Sleep.Disorder ~ BloodPressure_Upper + BloodPressure_Lower + Age + Stress.Level +Sleep.Duration+ Occupation +Heart.Rate
+ Daily.Steps
,  data = train_data, family = binomial())
summary(logistic_model_select)
vif(logistic_model_select)
pseudo_r2 <- pR2(logistic_model_select)
print(pseudo_r2)
predicted_probs <- predict(logistic_model_select, newdata=test_data,type = "response")
predicted_classes <- ifelse(predicted_probs > 0.4, 1, 0)
library(caret)
conf_matrix <- confusionMatrix(as.factor(predicted_classes), as.factor(test_data$Sleep.Disorder))
print(conf_matrix)
```

## logistic regression(手選變數 by 變數間相關係數/scatter plotej/共線性解決)
變數選取: BloodPressure_Upper + Stress.Level + Sleep.Duration +  BMI.Category  
```{r logist_self}
#BloodPressure_Upper + Stress.Level + Sleep.Duration +  BMI.Category   
logistic_model_original <- glm(Sleep.Disorder ~ BloodPressure_Upper + Stress.Level + Sleep.Duration +  BMI.Category ,  data = train_data, family = binomial())
summary(logistic_model_original)
library(car)
vif(logistic_model_original)
library(pscl)
pseudo_r2 <- pR2(logistic_model_original)
print(pseudo_r2)
predicted_probs <- predict(logistic_model_original,newdata=test_data, type = "response")
predicted_classes <- ifelse(predicted_probs > 0.4, 1, 0)
library(caret)
conf_matrix <- confusionMatrix(as.factor(predicted_classes), as.factor(test_data$Sleep.Disorder))
print(conf_matrix)
```


## random forest
```{r random_forest}
rf_model <- randomForest(Sleep.Disorder ~ Age + Gender + Occupation + Sleep.Duration + 
                         Quality.of.Sleep + Physical.Activity.Level + Stress.Level + 
                         BMI.Category + BloodPressure_Upper + BloodPressure_Lower + 
                         Heart.Rate + Daily.Steps, 
                         data = train_data, 
                         ntree = 500,  # Number of trees in the forest
                         mtry = 3,     # Number of predictors considered for each split
                         importance = TRUE)  # To calculate variable importance
print(rf_model)

predicted_classes <- predict(rf_model, newdata = test_data)
predicted_probabilities <- predict(rf_model, newdata = test_data, type = "prob")[, 2]

#  Model Evaluation 
# Confusion Matrix to assess performance
confusion_matrix <- confusionMatrix(predicted_classes, as.factor(test_data$Sleep.Disorder))
print(confusion_matrix)

# ROC Curve and AUC 
roc_curve <- roc(test_data$Sleep.Disorder, predicted_probabilities)
plot(roc_curve, main = "ROC Curve for Random Forest Model")
auc_value <- auc(roc_curve)
print(paste("AUC:", auc_value))

# Plot variable importance
var_imp <- importance(rf_model)
varImpPlot(rf_model, main = "Feature Importance in Random Forest")
```

## xgboost
```{r}
data_dummy <- model.matrix(Sleep.Disorder ~ ., data = data)[, -1] # Remove intercept
labels <- as.numeric(as.character(data$Sleep.Disorder)) # Target variable (0 or 1)

# Split the data into training and testing sets
set.seed(123) # For reproducibility
train_index <- createDataPartition(labels, p = 0.8, list = FALSE)
X_train <- data_dummy[train_index, ]
X_test <- data_dummy[-train_index, ]
y_train <- labels[train_index]
y_test <- labels[-train_index]
dtrain <- xgb.DMatrix(data = X_train, label = y_train)
dtest <- xgb.DMatrix(data = X_test, label = y_test)

# Set hyperparameters for the XGBoost model
param_list <- list(
  objective = "binary:logistic", # For binary classification
  eval_metric = "auc",           # We want to maximize AUC
  eta = 0.1,                     # Learning rate
  max_depth = 6,                 # Depth of the trees
  subsample = 0.8,               # Row sampling ratio
  colsample_bytree = 0.8,
  verbose = 1,                   # 訓練日誌詳細程度
  watchlist = list(train = dtrain, test = dtest), 
  early_stopping_rounds = 10# Feature sampling ratio
)

# Train the XGBoost model
set.seed(123)
xgb_model <- xgboost(
  data = dtrain, 
  params = param_list,          # Use params to specify objective
  nrounds = 100              # Print training log
#  watchlist = list(train = dtrain, test = dtest), 
 # early_stopping_rounds = 10  # Stop early if performance doesn't improve
)
# Predict probabilities on the test set
pred_probs <- predict(xgb_model, newdata = dtest)
# Convert probabilities to binary predictions (threshold = 0.5)
predictions <- ifelse(pred_probs > 0.5, 1, 0)
# Confusion matrix
confusion_matrix <- confusionMatrix(as.factor(predictions), as.factor(y_test))
print(confusion_matrix)
# Calculate AUC
auc <- roc(y_test, pred_probs)
print(auc$auc)

importance_matrix <- xgb.importance(model = xgb_model)
# Plot feature importance
xgb.plot.importance(importance_matrix)
```

## comparison
```{r comparison}
kable(data.frame(
  Metric = c('Accuracy', 
             'AUC', 
             'Multicollinearity', 
             'Feature Importance', 
             'Handles Nonlinearities', 
             'Computation Time'),
  XGBoost = c('Highest', 0.925, 'Not affected', 'Provides insights', 'Yes', 'Moderate'),
  Random_Forest = c('Higher', 0.913, 'Not affected', 'Provides insights', 'Yes', 'Slow'),
  Logistic_Regression = c('Lower', 0.889, 'Affected', 'Limited interpretability', 'No', 'Fast'))
)
```

## try cross validation
```{r cross}
train_control <- trainControl(
  method = "cv",  # k-fold cross-validation
  number = 10,    # Number of folds
)

#---------------------------------------所有變數

logist<-train(
  Sleep.Disorder ~ ., 
  data = data, 
  method = "glm",  # Specify "multinom" for multinomial logistic regression
  family = "binomial",       # Specify binary outcome
  trControl = train_control,
)
logist$results
print(logist)
# view final model 最終決定的模型，以及模型估計係數值
logist$finalModel
#view predictions for each fold，每一折(fold)/子集(subset)資料的預測誤差
logist$resample

#--------------------------------------------#stepwise變數

logist_step<-train(
  Sleep.Disorder ~ Sleep.Duration + Quality.of.Sleep + Physical.Activity.Level + Stress.Level + BloodPressure_Lower + Daily.Steps, 
  data = data, 
  method = "glm",  # Specify "multinom" for multinomial logistic regression
  family = "binomial",       # Specify binary outcome
  trControl = train_control,
)
logist_step$results
print(logist_step)
# view final model 最終決定的模型，以及模型估計係數值
logist_step$finalModel
#view predictions for each fold，每一折(fold)/子集(subset)資料的預測誤差
logist_step$resample

#------------------------------------elastic(還不確定)

# Define predictor variables
variablenames <- names(data)[-c(13)]  # Exclude unwanted columns
formula.x <- formula(paste("~", paste(variablenames, collapse=" + ")))
X <- model.matrix(formula.x, data)[, -1]  # Remove intercept column
y <- as.numeric(as.character(data$Sleep.Disorder))  # Ensure binary numeric target (0, 1)
table(y)
# Fit Elastic Net model with cross-validation
cv <- cv.glmnet(
  x = X, 
  y = y, 
  family = "binomial", 
  alpha = 0.5,        # Alpha controls the Elastic Net mixing (0: ridge, 1: LASSO)
  type.measure = "auc",  # Evaluate using AUC
  nfolds = 10         # Number of folds for cross-validation
)
# Extract coefficients for the best lambda (lambda.1se for simplicity)
coefs <- coef(cv, s = cv$lambda.1se)
# Print the best lambda
best_lambda <- cv$lambda.min
print(paste("Best lambda:", best_lambda))
# Extract non-zero coefficient variables (important features)
fre.variables <- rownames(coefs)[coefs[, 1] != 0]
fre.variables <- fre.variables[fre.variables != "(Intercept)"]  # Exclude intercept
print("Selected features:")
print(fre.variables)

#-----------------------------手選變數

logist_self<-train(
  Sleep.Disorder ~ BloodPressure_Upper + Stress.Level + Sleep.Duration +  BMI.Category, 
  data = data, 
  method = "glm",  # Specify "multinom" for multinomial logistic regression
  family = "binomial",       # Specify binary outcome
  trControl = train_control,
)
logist_self$results
print(logist_self)
# view final model 最終決定的模型，以及模型估計係數值
logist_self$finalModel
#view predictions for each fold，每一折(fold)/子集(subset)資料的預測誤差
logist_self$resample
```

## cross validation(repeated k-fold)
```{r}
train.rkfold <- trainControl(method = "repeatedcv", number = 5, repeats = 3)

logist1<-train(
  Sleep.Disorder ~ ., 
  data = data, 
  method = "glm",  # Specify "multinom" for multinomial logistic regression
  family = "binomial",       # Specify binary outcome
  trControl = train.rkfold,
)
print(logist1)
logist1$results
```









